---
title: "Kickstarter Data Exploration"
author: "Ian Scarff"
date: "January 24, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Data

Use the 2018 data file for kickstarters.

```{r}
data <- read.csv("ks-projects-201801.csv")
View(data)
```

#### Variables


* Name
    + The name of the kickstarter project
    
* Category
    + Name of sub-catagory under main catagories
    
* Main_Catagory
    + Name of main catagories.
    
* Currancy
    + The type of currancy used in pledging money (3 letter abbreviations).
    
* Deadline
    + Date kickstarter wants monetary goal to be reached (year - month - day).
    
* Goal
    + The monetary goal set by the kickstarter (in terms of type of currancy).
    
* Launched
    + Date the kick starter was started (year - month - day  hour:minute:second).
    
* Pledged
    + The amount of money given to the kickstarter (in terms of type of currancy).
    
* State
    + The current state of the kickstarter (Ex: successful, failed, ... etc.).
    
* Backers
    + The number if people who gave money.
    
* Country
    + The country of kickstarter origin (2 letter abbreviations).
    
* Usd.Pledged
    + The amount of money pledged in terms of the U.S. Dollar (rounded).
    
* Usd_Pledged_Real
    + The true amount of money (not rounded) pledged in terms of the U.S. Dollar.
    
* Usd_Goal_Real
    + The true monetary goal (not rounded) set by the kickstarter in terms of the U.S. Dollar.



#### Structure
```{r}
str(data)

```


The launched and deadline variables should be set as date variables.

```{r}
data$deadline <- as.Date(data$deadline)
data$launched <- as.Date(data$launched)

```

What is the time range for these projects?

```{r}
range(data$deadline)
range(data$launched)

```

Kickstarted was founded on April 28, 2009. Projects with launch dates in 1970 are not correct.

How many projects are there like this?

```{r}
which(data$launched == min(data$launched))
```

There are 7 such project. Remove them.

```{r}
data <- data[-c(which(data$launched == min(data$launched))), ]

range(data$launched)
```

##### Factor Variables

Let's look at some of the factor variables.

Main Categories
```{r}
levels(data$main_category)

```
There are 15 main categories

```{r}
table(data$main_category)
barplot(table(data$main_category), main = "Main Categories")

```

Film & Video is the largest category



Sub Categories

```{r}
levels(data$category)
```

There are 159 sub categories across the 15 main categories

```{r}
table(data$category)
barplot(table(data$category), main = "Sub Categories")


```

Product Design is the largest sub category



Currancy
```{r}
levels(data$currency)
```


There are 14 types of currancy

* AUD
    + Australian Dollar
    
* CAD
    + Canadian Dollar
    
* CHF
    + Swiss Frank
    
* DKK
    + Danich Krone
    
* EUR
    + Euro
    
* GBP
    + Great British Pound
    
* HKD
    + Hong Kong Dollar
    
* JPY
    + Japanese Yen
    
* MXN
    + Mexican Peso

* NOK
    + Norwegian Krone
    
* NZD
    + New Zealand Dollar
    
* SEK
    + Swedish Krona
    
* SGD
    + Singapore Dollar
    
* USD
    + United States Dollar
    
    
```{r}
### Make separate data frames for each currancy

aud <- length(which(data$currency == "AUD"))
cad <- length(which(data$currency == "CAD"))
chf <- length(which(data$currency == "CHF"))
dkk <- length(which(data$currency == "DKK"))
eur <- length(which(data$currency == "EUR"))
gbp <- length(which(data$currency == "GBP"))
hkd <- length(which(data$currency == "HKD"))
jpy <- length(which(data$currency == "JPY"))
mxn <- length(which(data$currency == "MXN"))
nok <- length(which(data$currency == "NOK"))
nzd <- length(which(data$currency == "NZD"))
sek <- length(which(data$currency == "SEK"))
sgd <- length(which(data$currency == "SGD"))
usd <- length(which(data$currency == "USD"))

```

```{r}
barplot(table(data$currency), main = "Currancy")
```

```{r}

### This function makes a pie graph
### Takes in the size of each slice, the labels for each slice, and the title
makePie <- function(slices, labels, title){

  ### Calculates what percent a slice takes of whole
  percent <- round(slices/sum(slices)*100)
  
  ### Creates the final label
  labels <- paste(labels,percent, sep = "  ")
  labels <- paste(labels,"%",sep="")
  labels <- paste(labels,slices, sep = " = ")
  
  ### Generates pie graph
  pie(slices, labels = labels, main = title)
  
  ### Displays labels
  labels
}


### Set slices and labels
CUR.SLICES <- c(aud,cad,chf,dkk,eur,gbp,hkd,jpy,mxn,nok,nzd,sek,sgd,usd)
CUR.LABELS <- levels(data$currency)

### Make pie graph
makePie(CUR.SLICES, CUR.LABELS, "Currency")




```

The United States Dollar is the most used currancy.



State

```{r}
levels(data$state)



```

* Canceled
    + A project was canceled before the deadline.
    
* Failed
    + A project did not make enough money by the deadline.
    
* Live
    + Projects that are currently on going.
    
* Successful
    + A project made or exceeded their goal by the deadline.

* Suspended
    + A project was suspended because it violated the terms & services of Kickstarter.
    
    
What is up with this undefined variable? Take a closer look.

```{r}
data.und <- data[data$state == "undefined",]
View(data.und)

head(data.und)
```

The rows with an undefined state level have many other problems. A big problem is the country. They are labeled under a country with the title of N,0", which is not a country. It can't be a misspelling of Norway because of the varying types of currancy throughout the rows.

```{r}
nrow(data.und)
```

There is a miniscule amount of rows that contain undefinded values. Remove them.

```{r}
data <- data[data$state != "undefined",]
table(data$state)

```

They have been removed.

```{r}

### Make separate data frames for each state
CANC <- length(which(data$state == "canceled"))
FAIL <- length(which(data$state == "failed"))
LIVE <- length(which(data$state == "live"))
SUCC <- length(which(data$state == "successful"))
SUSP <- length(which(data$state == "suspended"))

### Set slices and labels
ST.SLICES <- c(CANC, FAIL, LIVE, SUCC, SUSP)
ST.LABELS <- levels(data$state[-6])

### Make pie graph
makePie(ST.SLICES, ST.LABELS, "State")



```

Naturally, we expect the number of failures to be much greater than the number of successes. If we include canceled and suspended projects in the failed category, the number grows.


##### Statistics

Let's look at some statistics for the usd_goal_real variable.

```{r}
range(data$usd_goal_real)
mean(data$usd_goal_real)
median(data$usd_goal_real)

```

We see that the amount of money that projects request range wildly. The mean and median are drastically different, signifying there are large outliers and this variable is heavily right skewed.


Look at statistics for the usd_pledged_real variable.

```{r}
range(data$usd_pledged_real)
mean(data$usd_pledged_real)
median(data$usd_pledged_real)
```

We see that the amount of money that people pledge varies wildly. The mean and median are drastically different, signifying there are large outliers and this variable is heavily right skewed.


Look at statistics for the backers variable.

```{r}
range(data$backers)
mean(data$backers)
median(data$backers)
```
We see that the amount of backers for each project varies wildly. The mean and median are drastically different, signifying there are large outliers and this variable is heavily right skewed.



It is logical to think that most very large project would have either few backers or would have a high number of pledges and backers if successful. Most projects are smaller or mid-level in size.



##### New Variable

An important factor to think about is the amount of time the kickstarter runs for (i.e. the amount of time between the launch date and the deadline).

Create a new variable to hold these values.

```{r}
RunTime <- rep(NA, times = nrow(data))

### Calculate length of time
for (i in 1:length(RunTime)){
  
  RunTime[i] <- difftime(data$deadline[i], data$launched[i], units = "days")
  
}

```

Add this variable to the data

```{r}
data[16] <- RunTime
colnames(data)[16] <- "RunTime"
```



```{r}
range(data$RunTime)
mean(data$RunTime)
median(data$RunTime)
```

We can see that project run times vary, but have a similar mean and median.

```{r}
hist(data$RunTime)
abline(v = mean(data$RunTime), col = "red", lwd = 2, lty = 1)
abline(v = median(data$RunTime), col = "green", lwd = 2, lty = 2)
legend("topright", legend = c("mean","median"), col = c("red","green"), 
       lwd = c(2,2), lty = c(1,2))

```

We can see that there are some projects that take way longer than the mean or median, making the run time right skewd.



#### Next Step Considerations

For the purposes of this project, I propose we use only use kickstarters from the United States. Along with being the country with a large majority of the data set, it would remove undefined variable and that odd country name along with focusing on one currancy.

Since one of the statistical methods we will use is logistic regression, this eliminates country and currancy as predictors.

We still need to come up with another statistical methods to use and we need to form hypotheses to answer (Ex: If factor X important).



```{r}

data.US <- data[data$country == "US",]





```



























